---
title: "Activity: Visual Perception"
output: html_notebook
---

Purpose:

Create a set of figures that illustrate rungs on Cleveland and McGill's "visual hierarchy." These will be used in a survey that asks participants to judge numerical differences across groups.

# Setup

```{r}
library(tidyverse)
```


```{r}
df_mpg <- 
  mpg %>% 
  as_tibble()

df_mpg
```

## Pick variables

```{r}
df_mpg %>% 
  ggplot(aes(displ, color = class)) +
  geom_freqpoly(bins = 10) +
  geom_vline(
    data = . %>% 
      group_by(class) %>% 
      summarize(displ = median(displ)),
    mapping = aes(xintercept = displ, color = class)
  )
```

Observations
- The median `displ` has a good mix of differences. This will be our target for comparison across groups (by vehicle `class`).

## Create working dataset

```{r}
df_data <- 
  df_mpg %>% 
  group_by(class) %>% 
  summarize(displ = median(displ)) %>% 
  mutate(class = fct_reorder(class, displ))

df_data  
```

## Name values

For convenience, extract all values into the namespace:

```{r}
d_2seater <- 
  df_data %>% 
  filter(class == "2seater") %>% 
  pull(displ)

d_compact <- 
  df_data %>% 
  filter(class == "compact") %>% 
  pull(displ)

d_midsize <- 
  df_data %>% 
  filter(class == "midsize") %>% 
  pull(displ)

d_minivan <- 
  df_data %>% 
  filter(class == "minivan") %>% 
  pull(displ)

d_pickup <- 
  df_data %>% 
  filter(class == "pickup") %>% 
  pull(displ)

d_subcompact <- 
  df_data %>% 
  filter(class == "subcompact") %>% 
  pull(displ)

d_suv <- 
  df_data %>% 
  filter(class == "suv") %>% 
  pull(displ)
```


# Make Visuals

## 0. Demo

```{r}
tibble(
  category = c("A", "B"),
  value = c(1.0, 2.5)
) %>% 
  ggplot(aes(category, value)) +
  geom_col() +
  rzdr::theme_common()
ggsave("../images/fig0-demo.png")
```


## 1. Common scale

```{r}
df_data %>% 
  filter(class != "subcompact", class != "compact") %>% 
  
  ggplot(aes(class, displ)) +
  geom_col() +
  coord_flip() +
  rzdr::theme_common() +
  labs(
    x = "Class",
    y = "Engine Displacement (L)"
  )
ggsave("../images/fig1-common.png")
```

Pick some differences for tasks:

```{r}
# d_pickup - d_suv # 0.05, Too small
d_2seater - d_pickup  # 1.5, good
d_2seater - d_minivan # 2.9, good
d_2seater - d_midsize # 3.4, good

d_minivan - d_midsize # 0.5, good
```



## 2. Non-aligned scale

```{r}
df_data %>% 
  mutate(
    size = case_when(
      str_detect(class, "compact") ~ "small",
      class == "2seater" ~ "small",
      class %in% c("midsize", "minivan") ~ "midsize",
      class %in% c("suv", "pickup") ~ "utility"
    )
  ) %>% 
  
  ggplot(aes(class, displ)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~size, scales = "free_y") +
  rzdr::theme_common() +
  labs(
    x = "Class",
    y = "Engine Displacement (L)"
  )
ggsave("../images/fig2-nonaligned.png")
```

```{r}
d_midsize - d_subcompact # 0.6
d_pickup - d_minivan     # 1.4
d_pickup - d_compact     # 2.5

d_suv - d_compact        # 2.45, too many decimals
```


## 3. Length

```{r}
set.seed(102)

df_data %>% 
  filter(class != "subcompact", class != "compact") %>% 
  mutate(base = rnorm(n = n(), sd = 0.5)) %>% 
  bind_rows(tibble(
    class = "(Reference)",
    displ = 2,
    base = 2
  )) %>% 
  
  ggplot(aes(base, class, xend = base + displ, yend = class)) +
  geom_segment(size = 2) +
  annotate(
    "label",
    x = 4.5, y = "(Reference)",
    size = 5,
    label = "2 L",
  ) +
  
  scale_x_continuous(
    breaks = NULL,
  ) +
  rzdr::theme_common() +
  labs(
    x = "Engine Displacement (L)",
    y = "Class"
  )
ggsave("../images/fig3-length.png")
```

```{r}

```



## 4. Angle

```{r}
df_data %>% 
  mutate(
    size = case_when(
      str_detect(class, "compact") ~ "small",
      class == "2seater" ~ "small",
      class %in% c("midsize", "minivan") ~ "midsize",
      class %in% c("suv", "pickup") ~ "utility"
    )
  ) %>% 
  
  ggplot(aes(x = "", y = displ, fill = class)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  scale_y_continuous(
    # breaks = c(0, 2.5, 5, 7.5, 10),
    # labels = c("0", "2.5", "5", "7.5", "10")
    breaks = c(0, 5, 10),
    minor_breaks = NULL,
  ) +
  facet_wrap(~size) +
  rzdr::theme_common() +
  labs(
    x = "Class",
    y = "Engine Displacement (L)"
  )
ggsave("../images/fig4-angle.png")
```


## 5. Area

```{r}
df_data %>% 
  ggplot(aes(x = "", y = "")) +
  geom_point(aes(size = displ)) +
  facet_wrap(~class) +
  scale_size_area(max_size = 25) +
  theme_void() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),

    strip.text.x = element_text(size = 12),
    strip.text.y = element_text(size = 12),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90"),

    aspect.ratio = 4 / 4,

    plot.margin = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    plot.title = element_text(size = 18, hjust = 0),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 16, hjust = 1),
    plot.caption = element_text(size = 12, hjust = 1)
  ) +
  labs(x = "", y = "")
ggsave("../images/fig5-area.png")
```


## 6. Color saturation

```{r}
df_data %>% 
  ggplot(aes(x = "", y = "")) +
  geom_point(aes(color = displ), size = 25) +
  facet_wrap(~class) +
  viridis::scale_color_viridis() +
  theme_void() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),

    strip.text.x = element_text(size = 12),
    strip.text.y = element_text(size = 12),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90"),

    aspect.ratio = 4 / 4,

    plot.margin = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    plot.title = element_text(size = 18, hjust = 0),
    plot.title.position = "plot",
    plot.subtitle = element_text(size = 16, hjust = 1),
    plot.caption = element_text(size = 12, hjust = 1)
  ) +
  labs(x = "", y = "")
ggsave("../images/fig6-colorsat.png")
```


